Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-02-10 ribose-paper.RMD
tidied up rmd a bit

```{r nutrition analysis}
## Scripts
# - nut.analysis.R
# - dxa.R

#summarise macro
#nutritiontable with pro.pr kg bodyweight pr. group
#laod dxa data and take out weight and subject to join with nutrition data
fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
        select(weight, subject)

#nutrition data.
#the data is in grams and 
# added 50 grams protein as sup_protein for ingestion of whey protein isolat,
#And calculated in to the total amount of proteins ingested pr day.
#added 90g of glucose in to carbohydrate 
#added 360 kcal from glucose intake and calculated in total kcal pr day
#Load nutrition data.
nut.weight <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")%>%
        select(timepoint, meal, subject, protein, fat, calories,
               carbohydrates, group, sup_pro, sup_gluc, kcal_glu) %>%
        #changed names to compare groups each paired days
        mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
                                   "Day 1",
                                   if_else(timepoint %in% c("3", "4"),
                                           "Day 2",
                                           if_else(timepoint %in% c("5", "6"),
                                                   "Day 3",
                                                   if_else(timepoint %in% c("7", "8"),
                                                           "Day 4",
                                                           if_else(timepoint %in% c("9", "10"),
                                                                   "Day 5",
                                                                   if_else(timepoint %in% c("T3", "T4"),
                                                                           "Day 6",
                                                                           "na"))))))) %>%
        group_by(timepoint, subject, group) %>%
        #take in weight
        inner_join(fp.weight) %>%
        #summarise each variable to get total pr timepoint.
        #Whey protein supplement is added to total protein ingestion
        summarise(protein = sum(protein + sup_pro),
                  fat = sum(fat),
                  calories = sum(calories + kcal_glu),
                  carbohydrates = sum(carbohydrates + sup_gluc),
                  weight = sum(weight)) %>%
        #summarise to get protein pr weight.
        mutate(proprkg = (protein/weight)) %>%
        pivot_wider(names_from = group,
                    values_from = c(fat, protein, calories, carbohydrates, weight, proprkg))

#nutrition table with pro/weight. The table shows mean and sd of total intake 
#in grams pr. group, pr. timepoint.
#How much kcal on glucose supplement?
nuttable <- nut.weight %>%
        pivot_longer(names_to = "variable",
                     values_to = "values", col= fat_glucose:proprkg_placebo) %>%
        group_by(timepoint, variable) %>%
        summarise(m = mean(values, na.rm = TRUE),
                  s= sd(values, na.rm = TRUE)) %>%
        ungroup() %>%
        separate(variable, into = c("variable", "group")) %>%
        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
        select(timepoint, variable, group, stat,) %>%
        pivot_wider(names_from = variable,
                    values_from = stat) %>%
        select(timepoint, group, calories, carbohydrates, fat, protein,
               proprkg) %>%
        #changed names of group for simplicity 
        mutate(group = if_else(group %in% c("glucose"),
                               "G",
                               if_else(group %in% c("placebo"),
                                       "P",
                                       "")))

# Change data
#Does not contain FP 106 and 116 because of not delivered data.
#summarise macro
#Added sup_pro as whey protein isolate, 50 grams each group, each timepoint,
#90 grams glucose and 360 kcal.
fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
        select(weight, subject) 
nutha <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")%>%
        select(timepoint, subject, group, calories, fat, carbohydrates, protein, sup_pro, sup_gluc, kcal_glu) %>%
        mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
                                   "Day 1",
                                   if_else(timepoint %in% c("3", "4"),
                                           "Day 2",
                                           if_else(timepoint %in% c("5", "6"),
                                                   "Day 3",
                                                   if_else(timepoint %in% c("7", "8"),
                                                           "Day 4",
                                                           if_else(timepoint %in% c("9", "10"),
                                                                   "Day 5",
                                                                   if_else(timepoint %in% c("T3", "T4"),
                                                                           "Day 6",
                                                                           "na"))))))) %>%
        group_by(timepoint, subject, group) %>%
        inner_join(fp.weight) %>%
        summarise(protein = sum(protein + sup_pro),
                  fat = sum(fat),
                  calories = sum(calories + kcal_glu),
                  carbohydrates = sum(carbohydrates + sup_gluc),
                  weight = sum(weight)) %>%
        mutate(proprkg = (protein/weight))

#change data
#Protein
pro_change <- nutha %>%
        select(timepoint, subject, group, protein,) 
        #ungroup() %>%
        #pivot_wider(names_from = group,
         #           values_from = c(protein)) %>%
        #mutate(change = glucose - placebo) %>%
        #print()
prolm <- lmer(protein ~ timepoint + timepoint:group + (1|subject), data = pro_change)
pro.plot <- plot(prolm)
pro.sum <- summary(prolm)
#Fat
fat_change <- nutha %>%
        select(timepoint, subject, group, fat,) 
        #ungroup() %>%
        #pivot_wider(names_from = group,
         #           values_from = c(fat)) %>%
        #mutate(change = glucose - placebo) %>%
        #print()
fatlm <- lmer(fat  ~ timepoint + timepoint:group +(1|subject), data = fat_change)
fat.plot <- plot(fatlm)
fat.sum <- summary(fatlm)
#Carbohydrates
carb_change <- nutha %>%
        select(timepoint, subject, group, carbohydrates,) 
        #ungroup() %>%
        #pivot_wider(names_from = group,
         #           values_from = c(carbohydrates)) %>%
        #mutate(change = glucose - placebo) %>%
        #print()
carblm <- lmer(carbohydrates  ~ timepoint + timepoint:group + (1|subject), data = carb_change)
cho.plot <- plot(carblm)
cho.sum <- summary(carblm)
#calories
cal_change <- nutha %>%
        select(timepoint, subject, group, calories,) 
        #ungroup() %>%
        #pivot_wider(names_from = group,
         #           values_from = c(calories)) %>%
        #mutate(change = glucose - placebo) %>%
        #print()
callm <- lmer(calories  ~ 0 + timepoint + timepoint:group + (1|subject), data = cal_change)
cal.plot <- plot(callm)
cal.sum <- summary(callm)
#Protein pr bodyweigh
prowe <- nutha %>%
        select(timepoint, subject, group, proprkg) 
prokglm <- lmer(proprkg  ~ 0 + timepoint + timepoint:group + (1|subject), data = prowe)
prokglm.plot <- plot(prokglm)
prokglm.sum <- summary(prokglm)

```{r nutrition change analysis and fig}
# Associated scripts:
# - nut.change
# - nut.change.figs

>>>>

```{r nutrition change analysis and fig}
# Associated scripts:
# - nut.datahand.R
# - nut.change.R
# - nut.change.figs.R
